Aggregated Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2017-10-24 aggregation/agg-JIE.tex
basic aggregation; figure 4
More tweaks
private aggregation

@@

%\usepackage{lineno}
%\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

@@

\section{Formalizing LCA Studies}

@@

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, the use of a structured disclosure of study design introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model.  Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.

@@

\section{Protecting Privacy through Aggregation}

  \input{agg-agg}        % notation for full aggregation

%  \input{agg-pagg}       % notation for partial aggregation

%  \input{agg-forms}      % the elusive Figure 4

@@

%  \input{agg-review}     % review objectives

%  \input{agg-privacy}    % disclosure constraints

%  \input{agg-semantic}   % the need for stable references

@@

  %% \input{publication-foreground}

  %% \input{publication-example}

  %% \input{publication-discussion}

  %% \input{publication-conclusion}

  %% \input{publication-acknowledgments}

@@

  \input{publication-si-spreadsheet}

  \input{publication-forms}

  \input{publication-si-recipe}

  \input{publication-si-openaccess}

>>>>

\renewcommand{\topfraction}{0.7}

@@

%\usepackage{lineno}
\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

@@

\section{Disclosing the Product System Model}

@@

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, the use of a structured disclosure of study design introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model.  Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.  We return to the data privacy requirement in the next section.

@@

\section{Foreground Aggregation}

When an LCA study result is computed, the foreground and background matrices are aggregated into a scalar indicator score (or a small vector of a few scores).  The formulation in Eq.~\ref{eqn:study} and the illustration in Fig.~\ref{fig:foreground} provides a framework for reporting results at varying levels of aggregation.  In this section, I present several different aggregation computations that provide equivalent results for an LCA study.  After that, I show how a foreground containing private data can be partitioned into public and private portions, each of which can be comprehensively checked by a reviewer having privileged access, and subsequently disclosed in a manner that protects the private data.

  \input{agg-agg}         % the elusive Figure 4

  \input{agg-pagg}       % notation for partial aggregation

@@

 %% \input{agg-review}     % review objectives

 %% \input{agg-privacy}    % disclosure constraints

 %% \input{agg-semantic}   % the need for stable references

@@

%  \input{publication-foreground}

%  \input{publication-example}

%  \input{publication-discussion}

%  \input{publication-conclusion}

%  \input{publication-acknowledgments}

@@

%  \doublespacing

  \input{proofs}

%  \input{publication-si-spreadsheet}

%  \input{publication-forms}

%  \input{publication-si-recipe}

%  \input{publication-si-openaccess}

### 2017-10-24 aggregation/agg-lca-study.tex
private aggregation

It is currently common practice for background database maintainers to pre-compute the life cycle inventory results for their databases.  Using the notation $B_x = B \times (I - A)^{-1}$ to represent the pre-computed background LCI database, Eq.s~\ref{eqn:foreground} can be rewritten as:
\begin{equation}
 \tilde{A} = \left[\begin{array}{cc} 
A_f &  0 \\
A_d & 0 
   \end{array}\right];\;\;\;  \tilde{B} = \left[\begin{array}{cc} B_f & B_x \end{array}\right ]
\label{eqn:fg_bg}
\end{equation}
The proof of the equivalency of Eq.s~\ref{eqn:foreground} and \ref{eqn:fg_bg} is provided in the supplementary information to \cite{Kuczenski_JLCA_2015}.  In this form, the main content of the LCIDB has been removed from the $\tilde{A}$ matrix, and the only information remaining there pertains to the study foreground.  Furthermore, the final demand vector has been replaced with a canonical functional unit and no longer needs to be explicitly reported.

Using the ordered matrices, Eq.~\ref{eqn:leontief} can be written so that the computationally costly foreground and background inversions can be performed separately:\footnote{For a demonstration of this equivalence, the reader is again directed to \cite{Kuczenski_JLCA_2015}.}
\begin{equation}
s = \mathbf{e}^T \times (B_f + B_x\times A_d) \times (I - A_f)^{-1} \times \tilde{\mathbf{y}}
\label{eqn:study}
\end{equation}
Eq.~\ref{eqn:study} is visualized in Figure~\ref{fig:foreground}, with the background technology matrix $A$, having been subsumed into $B_x$, represented by an identity matrix $I$.

The activity levels of the foreground nodes are represented by $\tilde{\mathbf{x}}$, which is the result of $\tilde{\mathbf{y}}$ selecting the first column of $(I-A_f)^{-1}$.
%% An important intermediate result is the vector of activity levels for the foreground nodes $\tilde{\mathbf{x}}$, also visualized in Figure~\ref{fig:foreground}:
%% \begin{equation}
%% \tilde{\mathbf{x}} = (I - A_f)^{-1} \times \tilde{\mathbf{y}}
%% \label{eqn:inv}
%% \end{equation}
This vector can be used to describe an \emph{aggregated foreground}:
\begin{equation}
\begin{array}{rl}
    \tilde{\mathbf{b}}_f & = B_f \times \tilde{\mathbf{x}} \\
    \tilde{\mathbf{a}}_d & = A_d \times \tilde{\mathbf{x}} \\
    \tilde{\mathbf{b}}_x & = B_x \times \tilde{\mathbf{a}}_d
\end{array}
\label{eqn:agg}
\end{equation}
The study result can also be computed from the aggregated foreground:
\begin{equation}
\begin{array}{rl}
s & = \mathbf{e}^T \times( B_f + B_x \times A_d) \times \tilde{\mathbf{x}} \\
  & = \mathbf{e}^T \times( \tilde{\mathbf{b}}_f + B_x \times\tilde{\mathbf{a}}_d) \\
  & = \mathbf{e}^T \times( \tilde{\mathbf{b}}_f + \tilde{\mathbf{b}}_x) \\
\end{array}
\end{equation}

>>>>

It is currently common practice for background database maintainers to pre-compute the life cycle inventory results for their databases.  Using the notation $B_x = B \times (I - A)^{-1}$ to represent the pre-computed background LCI database, Eq.s~\ref{eqn:foreground} can be rewritten as:
\begin{equation}
 \tilde{A} = \left[\begin{array}{cc} 
A_f &  0 \\
A_d & 0 
   \end{array}\right];\;\;\;  \tilde{B} = \left[\begin{array}{cc} B_f & B_x \end{array}\right ]
\label{eqn:fg_bg}
\end{equation}
The proof of the equivalency of Eq.s~\ref{eqn:foreground} and \ref{eqn:fg_bg} is provided in the supplementary materials.  In this form, the main content of the LCIDB has been removed from the $\tilde{A}$ matrix, and the only information remaining there pertains to the study foreground.  Furthermore, the final demand vector has been replaced with a canonical functional unit and no longer needs to be explicitly reported.

Using the ordered matrices, Eq.~\ref{eqn:leontief} can be written so that the computationally costly foreground and background inversions can be performed separately:
\begin{equation}
s = \mathbf{e}^T \times (B_f + B_x\times A_d) \times (I - A_f)^{-1} \times \tilde{\mathbf{y}}_f
\label{eqn:study}
\end{equation}
where $\tilde{\mathbf{y}}_f$ is a canonical functional unit vector having the same rank as $A_f$. Eq.~\ref{eqn:study} is derived in the supplementary materials.
Eq.~\ref{eqn:study} is visualized in Figure~\ref{fig:foreground}, with the background technology matrix $A$, having been subsumed into $B_x$, represented by an identity matrix $I$.

The activity levels of the foreground nodes are represented by $\tilde{\mathbf{x}}$, which is the result of $\tilde{\mathbf{y}}_f$ selecting the first column of $(I-A_f)^{-1}$.
%% An important intermediate result is the vector of activity levels for the foreground nodes $\tilde{\mathbf{x}}$, also visualized in Figure~\ref{fig:foreground}:
%% \begin{equation}
%% \tilde{\mathbf{x}} = (I - A_f)^{-1} \times \tilde{\mathbf{y}}
%% \label{eqn:inv}
%% \end{equation}
This vector can be used to describe an \emph{aggregated foreground}:
\begin{equation}
\begin{array}{rl}
    \tilde{\mathbf{b}}_f & = B_f \times \tilde{\mathbf{x}} \\
    \tilde{\mathbf{a}}_d & = A_d \times \tilde{\mathbf{x}} \\
    \tilde{\mathbf{b}}_x & = B_x \times \tilde{\mathbf{a}}_d
\end{array}
\label{eqn:agg}
\end{equation}
The study result can also be computed from the aggregated foreground:
\begin{equation}
\begin{array}{rl}
s & = \mathbf{e}^T \times( B_f + B_x \times A_d) \times \tilde{\mathbf{x}} \\
  & = \mathbf{e}^T \times( \tilde{\mathbf{b}}_f + B_x \times\tilde{\mathbf{a}}_d) \\
  & = \mathbf{e}^T \times( \tilde{\mathbf{b}}_f + \tilde{\mathbf{b}}_x) \\
\end{array}
\end{equation}

### 2017-10-24 aggregation/agg-pagg.tex
private aggregation

\subsection{Aggregation and Privacy}

@@

>>>>

\subsection{Reviewable Private Aggregation}

@@

Often, a study author may wish to make a partial disclosure of a PSM that retains the privacy of confidential data, while still permitting comprehensive critical review of the complete model.  In this case it is necessary to grant a reviewer privileged access to the private data in order to meet the review objectives.  The study formulation in Eq.~\ref{eqn:study} may be used to partition a foreground into public and private portions in order to comply with any applicable constraints on what may be disclosed.  The approach is illustrated in Figure~\ref{fig:private}.

The first step is to identify the constraints on the disclosure.  In the framework presented here, constraints may include the locations or values of any subset of nonzero entries in the $A_d$ and $B_f$ matrices, as well as any subset of nodes in $A_f$.  Fig.~\ref{fig:private}a shows a PSM in which disclosure constraints are indicated by shaded regions of the matrices.  In this case the nodes represented by the three rightmost columns of $A_f$, as well as one entire row of $A_d$, are to remain secret.

The $A_d$ and $B_f$ matrices are then partitioned into two components that sum to the original:
\begin{equation}\begin{aligned}
  A_d =\; & A_{d,pub} + A_{d,priv} \\
  B_f =\; & B_{f,pub} + B_{f,priv}
  \label{eqn:partition}
  \end{aligned}
  \end{equation}
This is illustrated in Fig.~\ref{fig:private}b.  Substituted back into Eq.~\ref{eqn:study}, the private portions are aggregated into a private life cycle inventory vector $\mathbf{b}_{priv}$, while the public portions remain disaggregated:
\begin{align}
  s =\; & \mathbf{e}^T\times\left((B_{f,pub} + B_{f,priv}) + B_x\times (A_{d,pub} + A_{d,priv})\right)\times\tilde{\mathbf{x}}\\
  =\; & \mathbf{e}^T\times\left((B_{f,pub} + B_x\times A_{d,pub})\times\tilde{\mathbf{x}} + \mathbf{b}_{priv}\right)\\
  \mathbf{b}_{priv} =\; & (B_{f,priv} + B_x\times A_{d,priv})\times \tilde{\mathbf{x}}
\end{align}
Any foreground nodes that are completely contained within the private partition can be omitted from the disclosed foreground.  The aggregation result can then be included as a separate foreground node with a unit weight, and the aggregated inventory vector included as the corresponding column of $B_f$.  This is illustrated in Fig.~\ref{fig:private}c.

The reviewer must be granted access to both the public and private partitions in order to confirm that the partitioning is valid according to Eq.~\ref{eqn:partition}, and that $\mathbf{b}_{priv}$ was computed correctly; however, it is not thought possible for a reader to discern any information about $A_{d,priv}$ or $B_{f,priv}$ solely from $\mathbf{b}_{priv}$.  An advantage to this approach is that the reader can evaluate easily what fraction of the overall impact score is accounted for by the public versus the private portions of the model.  A ``disclosure completeness'' metric $\varphi$ can be defined as the fraction of the overall score which is accounted for by the disclosed portion of the PSM:
\begin{equation}
  \varphi =  \frac{\mathbf{e}^T\cdot(\mathbf{b} - \mathbf{b}_{priv})}{\mathbf{e}^T\cdot\mathbf{b}} = 1 - \frac{\mathbf{e}^T\cdot\mathbf{b}_{priv}}{\mathbf{e}^T\cdot\mathbf{b}}
  \label{eqn:metric}
\end{equation}
where $\mathbf{b}$ is the complete life cycle inventory.  The value of this metric can provide an indication of the level of transparency of the disclosure with respect to a given impact category indicator.  The author may relax the disclosure constraints by reporting the locations of some or all non-zero entries in $A_{d,priv}$ and $B_{f,priv}$ but not disclosing their values.

### 2017-10-24 aggregation/proofs.tex
private aggregation

We introduce the identity in between the two terms of the LCIDB:
\begin{align}
  \tilde{B}\times(I-\tilde{A})^{-1} &=
  \left[\begin{array}{cc}B_f & B\end{array}\right]\times I*  \times
  \left[\begin{array}{cc}I-A_f & 0 \\ -A_d & I-A\end{array}\right]^{-1}\\
      &=  \left[\begin{array}{cc}B_f & B\end{array}\right]\times 
      \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right]^{-1}\times
      \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right] \times
      \left[\begin{array}{cc}I-A_f & 0 \\ -A_d & Z\end{array}\right]^{-1}
\end{align}
Using the associative property of matrix multiplication, it can be shown that $M\times N^{-1} = (N\times M^{-1})^{-1}$ for any invertible $M$ and $N$ of the same rank, so:
\begin{eqnarray}
  \tilde{B}\times(I-\tilde{A})^{-1}
  &=& \left[\begin{array}{cc}B_f & B\end{array}\right]\times 
      \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right]^{-1}\times
\left(
\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & Z\end{array}\right]\times
    \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right]^{-1}\right)^{-1}\\
  &=&   \left[\begin{array}{cc}B_f & B\end{array}\right]\times 
      \left[\begin{array}{cc}I & 0 \\ 0 & Z^{-1}\end{array}\right]\times
\left(
\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & Z\end{array}\right]\times
\left[\begin{array}{cc}I & 0 \\ 0 & Z^{-1}\end{array}\right]\right)^{-1}\\
 &=& \left[\begin{array}{cc}B_f & B\times Z^{-1}\end{array}\right]\times
\left(\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & I\end{array}\right]\right)^{-1}
\end{eqnarray}
But these terms are equivalent to the flattened LCIDB:
\begin{eqnarray}
  \tilde{B}\times(I-\tilde{A})^{-1} &=&
  \left[\begin{array}{cc}B_f & B\times Z^{-1}\end{array}\right]\times
\left(\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & I\end{array}\right]\right)^{-1} \\
&=& \left[\begin{array}{cc}B_f & B\times (I-A)^{-1}\end{array}\right]\times
\left(I-\left[\begin{array}{cc}A_f & 0 \\ A_d & 0\end{array}\right]\right)^{-1}\\
&=& \tilde{B}_{flat}\times(I-\tilde{A}_{flat})^{-1}
\end{eqnarray}
thus completing the proof.

>>>>

We introduce the identity in between the two terms of the LCIDB:
\begin{align}
  \tilde{B}\times(I-\tilde{A})^{-1} &=
  \left[\begin{array}{cc}B_f & B\end{array}\right]\times I*  \times
  \left[\begin{array}{cc}I-A_f & 0 \\ -A_d & I-A\end{array}\right]^{-1}\\
      &=  \left[\begin{array}{cc}B_f & B\end{array}\right]\times 
      \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right]^{-1}\times
      \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right] \times
      \left[\begin{array}{cc}I-A_f & 0 \\ -A_d & Z\end{array}\right]^{-1}
\end{align}
Using the associative property of matrix multiplication, it can be shown that $M\times N^{-1} = (N\times M^{-1})^{-1}$ for any invertible $M$ and $N$ of the same rank, so:
\begin{eqnarray}
  \tilde{B}\times(I-\tilde{A})^{-1}
  &=& \left[\begin{array}{cc}B_f & B\end{array}\right]\times 
      \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right]^{-1}\times
\left(
\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & Z\end{array}\right]\times
    \left[\begin{array}{cc}I & 0 \\ 0 & Z\end{array}\right]^{-1}\right)^{-1}\\
  &=&   \left[\begin{array}{cc}B_f & B\end{array}\right]\times 
      \left[\begin{array}{cc}I & 0 \\ 0 & Z^{-1}\end{array}\right]\times
\left(
\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & Z\end{array}\right]\times
\left[\begin{array}{cc}I & 0 \\ 0 & Z^{-1}\end{array}\right]\right)^{-1}\\
 &=& \left[\begin{array}{cc}B_f & B\times Z^{-1}\end{array}\right]\times
\left(\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & I\end{array}\right]\right)^{-1}
\end{eqnarray}
But these terms are equivalent to the flattened LCIDB:
\begin{eqnarray}
  \tilde{B}\times(I-\tilde{A})^{-1} &=&
  \left[\begin{array}{cc}B_f & B\times (I-A)^{-1}\end{array}\right]\times
\left(\left[\begin{array}{cc}I-A_f & 0 \\ -A_d & I\end{array}\right]\right)^{-1} \\
&=& \left[\begin{array}{cc}B_f & B_x\end{array}\right]\times
\left(I-\left[\begin{array}{cc}A_f & 0 \\ A_d & 0\end{array}\right]\right)^{-1}\\
&=& \tilde{B}_{flat}\times(I-\tilde{A}_{flat})^{-1}
\end{eqnarray}
thus completing the proof.

\subsection{The Study Foreground Equation}

Equation 5 in the manuscript is derived from Equation 4.

We begin by computing $(I-\tilde{A})^{-1}$, again taking advantage of the matrix's block triangularity.  Find $Q$ such that:
\begin{equation}
  \left[\begin{array}{cc} (I-A_f) & 0 \\ -A_d & I\end{array}\right]\times
  \left[\begin{array}{cc} (I-A_f)^{-1} & 0 \\ Q & I\end{array}\right] =
  \left[\begin{array}{cc} I & 0 \\ 0 & I\end{array}\right]
\end{equation}

It can be seen that $Q=A_d\times(I-A_f)^{-1}$ is a necessary condition to satisfy the equality, and thus:
\begin{equation}
  (I-\tilde{A})^{-1} = \left[\begin{array}{cc} (I-A_f)^{-1} & 0 \\ A_d\times(I-A_f)^{-1} & I\end{array}\right]
\end{equation}.

Substituting this into the LCA system equation:
\begin{align}
  s =\;& \mathbf{e}^T\times \tilde{B}\times(I-\tilde{A})^{-1}\tilde{\mathbf{y}}\\
  =\;& \mathbf{e}^T\times \left[\begin{array}{cc} B_f & B_x \end{array}\right]\times
  \left[\begin{array}{cc} (I-A_f)^{-1} & 0 \\ A_d\times(I-A_f)^{-1} & I \end{array}\right]
  \times\left[\begin{array}{c} \tilde{\mathbf{y}}_f \\ \mathbf{0}\end{array}\right]
\end{align}
where $\tilde{\mathbf{y}}_f$ is a canonical functional unit having the same dimension as $A_f$.  Thus:
\begin{align}
s =\;& \mathbf{e}^T\times \left[\begin{array}{cc} B_f & B_x \end{array}\right]\times
 \left[\begin{array}{c} (I-A_f)^{-1}\times\tilde{\mathbf{y}}_f \\ A_d\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f\end{array}\right]\\
 =\;& \mathbf{e}^T\times \left(B_f\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f
 + B_x\times A_d\times(I-A_f)^{-1}\times\tilde{\mathbf{y}}_f\right)\\
 =\;& \mathbf{e}^T\times \left(B_f + B_x\times A_d\right)\times(I-A_f)^{-1}\times \tilde{\mathbf{y}}_f\\
 =\;& \mathbf{e}^T\times \left(B_f + B_x\times A_d\right)\times \tilde{\mathbf{x}}
\end{align}
thus completing the derivation.
